---
title: "Limitations of fitting power law to secondary titers"
subtitle: "Visualizing error metrics for fitting power law decay to observed titers after a secondary response"
format: html
---

```{r setup}
# loading libraries
library(here) #for referencing relative to .rproject repo
library(ggplot2) #for plotting
library(tidyverse) #for easy manipulating dataframes
library(lme4) 
library(gridExtra)

# code from source
source(here("code", "functions", "functions.R"))
```

# Objective

The purpose of this document is to visualize how the error in power law fits to observed titers after secondary repsonses changes with the time between responses ($\tau$) and the fold change in observed titers after the boost ($\delta$).

We have previously calculated the AIC for power law fits, the estimated decay rate, and the error in the estimated decay rate from the true decay rate. Here, we visualize these quanttites as heat maps.

# Load Analysis Results

Below, we load the AIC for power law fits, the estimated decay rate, and the error in the estimated decay rate from the true decay rate. 

```{r load-analysis-results}
#AICs
AICs = read.csv(here("results", "tables", "analysis", "AICs.csv"))

#k estimates
estimated.ks = read.csv(here("results", "tables", "analysis", "estimated_ks.csv"))

#error in k 
error.k = read.csv(here("results", "tables", "analysis", "error_k.csv"))

# load taus and deltas
tau.and.delta = read.csv(here("data", "parameters", "tau_and_delta.csv"))

```

To plot these as base r contour plots, we need to reformat them as matrices with no $\tau$ column

```{r reformat}
# add column names to indicate deltas
# extracting taus and deltas
taus = unique(tau.and.delta$tau)
deltas = unique(tau.and.delta$delta)

#formating dataframes as matrices with no labels
AICs = AICs[,-1] %>% as.matrix()
estimated.ks = estimated.ks[,-1] %>% as.matrix()
error.k = error.k[,-1] %>% as.matrix()
```

below, we plot these as contour plots: 

```{r contour-plots}
#plot AICs
filled.contour(x = taus, y = deltas, z=AICs, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "AIC of Power Law Fit to Observed Secondary Response Titers")

#plot estimated ks
filled.contour(x = taus, y = deltas, z=estimated.ks, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "Estimated k of Power Law \n Fit to Observed Secondary Response Titers")

#plot estimated ks
filled.contour(x = taus, y = deltas, z=error.k, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "Fractional Error of k of Power Law \n Fit to Observed Secondary Response Titers")

########### save figures ##################
file = here("results", "figures", "fit limitations", "AICs.png") #name file path
png(file) #open developer
  filled.contour(x = taus, y = deltas, z=AICs, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "AIC of Power Law Fit to Observed Secondary Response Titers") #draw figure
dev.off() #close developer

file = here("results", "figures", "fit limitations", "estimated_ks.png")
png(file)
 filled.contour(x = taus, y = deltas, z=estimated.ks, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "Estimated k of Power Law \n Fit to Observed Secondary Response Titers")
dev.off()

file = here("results", "figures", "fit limitations", "error_k.png")
png(file)
  filled.contour(x = taus, y = deltas, z=error.k, lwd = 2, xlab = expression(tau), ylab = expression(delta), main = "Fractional Error of k of Power Law \n Fit to Observed Secondary Response Titers")
dev.off()
```


# heatmaps with ggplot

To plot these as heatmaps, we need to format them in long form. First, I reload the data:

```{r reload-analysis-results}
#AICs
AICs = read.csv(here("results", "tables", "analysis", "AICs.csv"))

#k estimates
estimated.ks = read.csv(here("results", "tables", "analysis", "estimated_ks.csv"))

#error in k 
error.k = read.csv(here("results", "tables", "analysis", "error_k.csv"))

```

```{r format-long-form}
# add column names to indicate deltas
# extracting taus and deltas
taus = unique(tau.and.delta$tau)
deltas = unique(tau.and.delta$delta)

# adding row to indicate tau and columns to indicate delta
colnames(AICs) = c("tau", deltas) 
colnames(estimated.ks) = c("tau", deltas) 
colnames(error.k) = c("tau", deltas) 

#AIC 
AICs = AICs %>% pivot_longer(cols = !tau,
                             names_to = "delta",
                             values_to = "AIC")
#Estimated K
estimated.ks = estimated.ks %>% pivot_longer(cols = !tau,
                                             names_to = "delta",
                                             values_to = "estimated.k")

#fractional error
error.k = error.k %>% pivot_longer(cols = !tau,
                                   names_to = "delta",
                                   values_to = "error.k")

```

# Plotting Heatmaps 

Below, we plot heatmaps of AIC, estimated $k$, and fractional error in estimated $k$.

```{r generate-heatmaps}
#AICs
plot1 = ggplot() + geom_tile(data = AICs, aes(x = tau, y = delta, fill = AIC)) +
  theme_bw() + 
  labs(fill = "AIC", title = "")
plot1

#Estimated Ks
plot2 = ggplot() + geom_tile(data = estimated.ks, aes(x = tau, y = delta, fill = estimated.k)) +
  theme_bw() +
  labs(fill = "k estimate")
plot2

#Percent Error in K
plot3 = ggplot() + geom_tile(data = error.k, aes(x = tau, y = delta, fill = error.k)) +
  theme_bw() +
  labs(fill = "fractional error in k from true")
plot3
```
