---
title: "Secondary Responses"
subtitle: "Generating a time series of power-law decaying antibody titers after secondary vaccination or infection"
format: html
---

```{r setup}
# loading libraries
library(here) #for referencing relative to .rproject repo
library(ggplot2) #for plotting
library(dplyr) #for easy manipulating dataframes
library(lme4) 

# code from source
source(here("Theory", "code", "functions", "functions.R"))
```

# Objective

The goal of this document is to generate a time series of antibody titer decay according to a power law post secondary response. 

We will generate time series of secondary responses varying the time between primary and secondary response ($\tau$) and the fold change at the time of secondary response ($\delta$).

# Equations

Antibody titers generated after a primary response decay according to a power law that begins at $Time_0$ and peaks at $Time=1$ with titer $Ab_1$: 

$$
Ab(Time) = AB_1(Time)^k
$$

The time between primary and secondary response is $\tau$. The fold change in titers at the peak of the secondary response ($time = \tau + 1$) is $\delta$. This fold change is the sum of the titer remaining from the primary response ($Ab_1(\tau+1)^k$) and the newly generated titer.

The newly generated titers decay according to the power law $Ab(Time) = (\delta Ab_1(\tau)^k - Ab_1(\tau+1)^k)(Time - \tau)^k$. 

The overall observed anitbody titer will be the sum of these two populations: 

$$
Ab(Time) = Ab_1(\tau+1)^k + (\delta Ab_1(\tau)^k - Ab_1(\tau+1)^k)(Time - \tau)^k
$$
Ultimately, we would like to know how $\delta$ and $\tau$ effect the fit of a power law model to this non-power law sum, both in terms of AIC and in terms of the error in the estimate of the true decay parameter, $k$.

In this document, we will vary $\delta$ and $\tau$ to generate time series of a secondary response. We will save these time series to a data frame for model fitting. 

# Generating secondary responses

In generating our time series of primary and secondary antibody repsonses, we would like to track: 

1) Titers remaining from primary response (follows a true power law);
2) Titers that are newly generated (follows a true power law); and
3) The observed sum of these two titers (does not follow a true power law)

## Setting parameters 

First, we generate a set of combinations of $\tau$ and $\delta$. 

```{r generate-parms-df}
# set taus in terms of months - leaving at least 3 years for second boost time series
taus = seq(1,17)*12 # time in months, so seq is in years and multiply by 12 for months
# set deltas
deltas = seq(2,10, by = 0.5)

# generate combination matrix and set column names
parameters = expand.grid("tau" = taus, "delta" = deltas, KEEP.OUT.ATTRS = F) # should have length(deltas) x length(taus) rows; keep out attributes = F stops the return of metadata

# save parameters df 
file = here("Theory", "data", "parameters", "tau_and_delta.csv")
write.csv(parameters, file = file, row.names = F)
```

Next, we load our parameters for the power law decay after primary response.

```{r load-power-law-parms}
# load saved parms from csv
power.law.parms = read.csv(here("Theory", "data", "parameters", "power_law_parms.csv"))

# save individually to environment
A1 = power.law.parms[1, "A1"] #peak of primary response
k = power.law.parms[1, "k"] #power law decay constant
Tmax = power.law.parms[1, "Tmax"] #maximum time for both responses
```

## Generating time series

We will iterate over all rows of the $\tau$ and $\delta$ combinations dataframe generated in the code chunk above to simulate a secondary response. For each secondary response, we will save both the titers from response 1 and response 2, as well as the sum of both, from $T_0$ of the primary response to $Tmax$, in a dataframe. 

```{r generate-secondary-responses}
# iterate over parameter df for every combination of tau and delta upon secondary response
time.series = lapply(t(parameters) %>% as.data.frame(), function(parms){ #transpose because lapply takes rowwise
  ## set times for second response , tau, and delta
  tau =  parms[1] #from parms df
  delta = parms[2] #from parms df
  times = seq(tau+1, Tmax, by = 0.5)
  
  ## generate antibody titers from primary and secondary responses at time tau + 1 to Tmax
  primary.response = power_law_function(A1 = A1, k = k, time = times) #remaining primary response
  secondary.response = additive_power_law_function(A1 = A1, k = k, tau = tau, delta = delta, time = times) # new secondary response 
  observed = primary.response + secondary.response #observed titer is the sum of both
  
  ## format dataframe of both populations and observed, add time and label with tau and delta
  boost = data.frame("response" = 2, "tau" = tau, "delta" = delta, "time" = times, "response.1.titer" = primary.response, "response.2.titer" = secondary.response, "observed" = observed)
  
  ## add titers from time 0 to time tau
  primary.pre.tau = read.csv(here("Theory", "data", "primary response", "primary_response.csv")) #load primary titers
  primary.pre.tau = primary.pre.tau %>%
    filter(time <= tau) %>%
    mutate("response.2.titer" = 0,
           "observed" = response.1.titer + response.2.titer,
           "tau" = tau,
           "delta" = delta)
  
  ## bind pre tau and post tau titer dfs
  time.series = rbind(primary.pre.tau, boost) %>% as.data.frame()
  
  return(time.series)
})

## concatenated all dfs into a list to easily save
time.series = do.call(rbind, time.series) %>% as.data.frame()

## save to data
file = here("Theory", "data", "secondary response", "secondary_responses.csv")
write.csv(time.series, file = file, row.names = F)
```

Now, we have a dataframe of time series of primary and secondar responses from each combination of $\tau$ and $\delta$.

We will visualize a few of these in the file `code\generating_data\3_response_figures.qmd`. 
We will fit the observed titers after the secondary response to a power law in `code\analysing_data\4_model_fitting.qmd`.
