---
title: "Visualizing Secondary Responses"
subtitle: "Plotting example time series of primary and secondary responses"
format: html
---

```{r setup}
# loading libraries
library(here) #for referencing relative to .rproject repo
library(ggplot2) #for plotting
library(tidyverse) #for easy manipulating dataframes
library(lme4) 
library(gridExtra)

# code from source
source(here("Theory", "code", "functions", "functions.R"))
```

# Objective

The goal of this document is to visualize titers generated after primary and secondary responses. We will plot: 

1) titers generated from the primary response,
2) titers generated from the secondary response, and
3) what we observe, the sum of these two. 

The difference in the observed titers from the secondary titers alone will depend on the time between primary and secondary resposnes, $\tau$, and the magnitude of the fold change on the boost, $\delta$. 

We have simulated secondary responses for many combinations of $\tau$ and $\delta$. Here, we choose a few demonstrative combinations.

# Loading parameter values and data

The parameters for the power law decay of primary titers are saved to `data/parameters/power_law_parms.csv`. 

The combinations of $\tau$ and $\delta$ used to generated secondary responses are saved to `data/parameters/tau_and_delta.csv`. 

The time series of primary and secondary responses are saved to `data/secondary response/secondary_responses.csv`.

We load these below.

```{r load-parameters}
# power law parms
power.law.parms = read.csv(here("Theory", "data", "parameters", "power_law_parms.csv"))

# secondary responses parms
delta.and.tau = read.csv(here("Theory", "data", "parameters", "tau_and_delta.csv"))

# time series
time.series = read.csv(here("Theory", "data", "secondary response", "secondary_responses.csv"))
```

# Visualizing 

We would like to plot 4 different time series: 

1) $\tau$ is short and $\delta$ is small;
2) $\tau$ is long and $\delta$ is small;
3) $\tau$ is long and $\delta$ is large; and 
4) $\tau$ is short and $\delta$ is large.

We choose these four combinations below: 

```{r set-demonstrative-tau-and-delta}
# small and large tau
tau.small = range(delta.and.tau$tau)[1]
tau.large = range(delta.and.tau$tau)[2]

# small and large delta
delta.small = range(delta.and.tau$delta)[1]
delta.large = range(delta.and.tau$delta)[2]

# create a list of all combinations
combinations = expand.grid("tau" = c(tau.small, tau.large), "delta" = c(delta.small, delta.large), KEEP.OUT.ATTRS = F) #four combinations means four figures
combinations = t(combinations) %>% as.data.frame() #transpose for columnwise use in lapply
```

We filter our time series for only these unique combinations of $\tau$ and $\delta$. Below, we plot each time series from $Time_0$ of the primary response to $Tmax$, the total time of observing titers.

```{r plot-titers}
# filter and plot each unique combo of tau and delta; save plots in list
plots = lapply(combinations, function(combination){
  tau = combination[1]
  delta = combination[2]
  # set combination as df
  scenario = data.frame("tau" = tau, "delta" = delta) #restructure as df for use in join
 
  # semi join so only rows that match tau and delta combination are kept
  time.series.filtered = semi_join(time.series, scenario) %>%
    select(!c(response,tau,delta)) %>%
    pivot_longer(cols = !time, 
                 names_to = "response",
                 values_to = "titer") %>% as.data.frame()
 
  # plot
  plot = ggplot() + 
    #plot populations
    geom_line(data = time.series.filtered %>% 
                filter(response == "observed"), 
              aes(x = time, y = log10(titer), linetype = "observed")) +
     geom_line(data = time.series.filtered %>% 
                filter(response == "response.1.titer") %>% 
                filter(time > tau), 
              aes(x = time, y = log10(titer), linetype = "primary response")) + 
    geom_line(data = time.series.filtered %>% 
                filter(response == "response.2.titer"),
                aes(x = time, y = log10(titer), linetype = "secondary response")) +
    #labels 
    labs(x = "Months", y = "Log10 Titer", title = paste("Tau is", tau, "and delta is", delta), linetype = "") +
    #aesthetics
    theme_bw() 
})

# plot all to a single figure
do.call(grid.arrange, plots)

# Save figure 
file = here("Theory", "results", "figures", "secondary response dynamics", "example_time_series.png")
ggsave(do.call(grid.arrange, plots), file = file)
```


# Poster figure 

I would like to show two different regimes - one where the error in the power law estimate is low, and one where the error in the power law estimate is high. 

```{r set-tau-delta}
# tauu = 10 and delta = 3 vs tau = 5 and delta = 5
tau.low = 5 #in years
delta.low = 10
tau.high = 10
delta.high = 5

# poor fit. tau high and delta low, good fit - tau low and delta high
combinations = data.frame("tau" = c(tau.low, tau.high), "delta" = c(delta.high, delta.low), "fit" = c("good", "bad")) %>% t(.) %>% as.data.frame()
```

```{r generate-plots}
# filter and plot each unique combo of tau and delta; save plots in list
plots = lapply(combinations, function(combination){
  tau = as.numeric(combination[1])*12 #in months
  delta = as.numeric(combination[2])
  fit = combination[3]
  # set combination as df
  scenario = data.frame("tau" = tau, "delta" = delta) #restructure as df for use in join
 
  # semi join so only rows that match tau and delta combination are kept
  time.series.filtered = semi_join(time.series, scenario) %>%
    select(!c(response,tau,delta)) %>%
    pivot_longer(cols = !time, 
                 names_to = "response",
                 values_to = "titer") %>% as.data.frame()
 
  # plot
  plot = ggplot() + 
    #plot populations
    geom_line(data = time.series.filtered %>% 
                filter(response == "observed"), 
              aes(x = time/12, y = log10(titer), linetype = "observed")) +
     geom_line(data = time.series.filtered %>% 
                filter(response == "observed",
                       time > tau) , 
              aes(x = time/12, y = log10(titer), linetype = "observed", col = fit)) +
     geom_line(data = time.series.filtered %>% 
                filter(response == "response.1.titer") %>% 
                filter(time > tau), 
              aes(x = time/12, y = log10(titer), linetype = "primary response")) + 
    geom_line(data = time.series.filtered %>% 
                filter(response == "response.2.titer"),
                aes(x = time/12, y = log10(titer), linetype = "secondary response")) +
    #labels 
    labs(x = "Years", y = "Log10 Titer", title = paste("Tau is", tau/12, "and delta is", delta), linetype = "") +
    #aesthetics
    theme_bw() +
       scale_color_manual(values = c("good" = "goldenrod", "bad" = "orangered3")) +
       guides(color = "none")
})

# plot all to a single figure
do.call(grid.arrange, plots)

# Save figure 
file = here("Theory", "results", "figures", "fit limitations", "example_time_series.png")
ggsave(do.call(grid.arrange, plots), file = file)


```

# Next steps

We will now fit the observed time series on secondary response to a power law, and plot how $\tau$ and $\delta$ affect the $AIC$ and accuracy of the underlying power law constant, $k$. We will perform this analysis in the documents in `code/analysing_data`.
