---
title: "First Response"
subtitle: "Generating a time series of power-law decaying antibody titers after primary vaccination or infection"
format: html
---

```{r setup}
# loading libraries
library(here) #for referencing relative to .rproject repo
library(ggplot2) #for plotting
library(dplyr) #for easy manipulating dataframes
library(lme4) 

# code from source
source(here("code", "functions", "functions.R"))
```

# Objective 

The purpose of this document is to generate a time series of antibody titer decay according to a power law post primary immune response. The magnitude of the response and power law decay coefficient will be the same for all subsequent simulated secondary infections; the newly generated titers of the secondary response will boost and decay with a magnitude proportional to the magnitude of the primary response and the power law decay coefficient.

# Setting parameters

We will define 3 parameters: the magnitude of a primary boost in antibody titers, $A1$; the power law decay constant, $k$, which is less than 0; and the total time frame over the course of which both a primary and secondary response occurs.

```{r static-parameters}
# True power law parameters
A1 = 10^5 # peak antibody titer at time = 1 month
k = -0.5 #antibody decay constant

#Tmax for simulations
Tmax = 20 #years to simulate both responses over
Tmax = Tmax*12 # number of months to simulate responses over (time in months)

# save parameter values to df
power.law.parms = data.frame("A1" = A1, "k" = k, "Tmax" = Tmax) #safe as df

# save df as csv 
file = here("data", "parameters", "power_law_parms.csv") #filepath
write.csv(power.law.parms, file = file, row.names = F) #save to filepath
```

DONT FORGET TO SAVE THESE AS A TABLE

# Generating primary response time series

We assume that $Time = 0$ is the time of the primary infection or vaccination. We take the peak of the response to occur ~30 days after time = 0, and adjust our time scale such that 1 unit time = 30 days. In other words, the power law parameter $A1$ will be the anitbdy titer at the peak of response. 

In the code chunk below, we generate a time series according to a power law. 

```{r generate-primary-decaying-titers}
### set time points for titer collection
times = seq(1, Tmax, by = 0.5) #every two weeks 

### calculate titers
titers = sapply(times, FUN = function(time) {(power_law_function(A1 = A1, k = k, time = time))})

# format as data frame
primary.response = data.frame("response" = 1, "time" = times, "response.1.titer" = titers)
primary.response = rbind(data.frame("response" = 1, "time" = 0, "response.1.titer" = 0), primary.response) %>% as.data.frame() #add titer at T=0

# save to data 
file = here("data", "primary response", "primary_response.csv") #filepath
write.csv(primary.response, file = file, row.names = F) #save to filepath
```

# Visualize

Let's plot these titers from the time of primary exposure ($time = 0$) to the total time of both responses, ($time = Tmax$).

```{r plot-primary-response-titers}
# plot
plot = ggplot() + geom_point(data = primary.response, aes(x = time, y = response.1.titer)) + 
  geom_line(data = primary.response, aes(x = time, y = response.1.titer, linetype = "Observations")) +
  # plot power law 
   stat_function(fun = function(x){power_law_function(A1 = A1, k = k, time = x)}, aes(linetype = "Power Law")) +
  #aesthetics
  theme_bw() +
  labs(x = "Months", y = "Ab titer", title = "Primary Response Synthetic Data") +
  scale_linetype_manual(values = c("Observations" = "solid", "Power Law" = "dashed"))
 

#print to plot viewer
print(plot)

# save to results
file = here("results", "figures", "primary response dynamics", "primary_response.png")
ggsave(plot = plot, filename = file)
```

The solid line represents observations generated according to a power law, and decay begins at $time = month$. We chose one month as our time step because the power law model 

$$ 
Ab(Time) = Ab_1(Time)^{k} , \text{ where } k<0
$$
takes the titer at $time = 1$ unit after power law decay begins ($Time_0$) as a parameter, $Ab_1$. This model takes $Ab_1$ to be the antibody titer at the peak of the response, which roughly occurs one month after infection. 


Note that "when" power law decay in antibody titers starts, that is, $Time_0$, is an open question, and one that forces us to confront whether the power law is phenomenological or a true result of the mechanism of plasma cell decay.
